Parallel Monte Carlo Synthetic Acceleration Methods for Discrete Transport Problems

### 2013-09-03 doc/dissertation/presentation/defense.tex
restructuring nonlinear results

  \begin{itemize}
  \item Inject MCSA as the solver at each eigenvalue operation
    \medskip
  \item Transport operator is static - one time cost for MCSA setup
    \medskip
  \item Current Exnihilo implementation gives the full operator
  \end{itemize}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=1.9in]{solver_time.pdf}
        \end{center}
      \end{figure}

@@

      \begin{itemize}
        {\small
        \item Comparison to Trilinos Aztec Krylov solvers with ILUT
          \medskip
        \item MCLS generates the same k-eigenvalue and neutron flux in
          all groups and spatial locations
          \medskip
        \item MCSA converged in fewer iterations than GMRES, more
          iterations than BiCGStab
          \medskip
        \item Explicit preconditioning strategy destroys sparsity and
          elevates CPU times (Ifpack ILUT)
          \medskip
        \item Spectral radius and memory limitations combine to
          prevent solutions at finer discretizations
        }
      \end{itemize}

@@

  \begin{itemize}
  \item Sequence of Navier-Stokes benchmarks
    \begin{itemize}
    \item Thermal convection cavity problem (De Vahl Davis, 1983)
    \item Lid driven cavity problem (Ghia et al., 1982)
    \item Backward-Facing step problem (Gartling, 1990)
    \end{itemize}
    \medskip
  \item Tuning benchmark parameters varies the strength of
    nonlinearities
    \medskip
  \end{itemize}

@@

  \begin{itemize}
  \item Need a nonlinear solution scheme for the Navier-Stokes equations
    \bigskip
  \item Significant research on Newton methods since the 1980's
    \bigskip
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Robust implementations
    \item No operator required
    \end{itemize}
    \bigskip
  \item Monte Carlo methods need the full operator
    \bigskip
  \item Automatic construction of the linear model is available
    \begin{itemize}
    \item Operator overloading for nonlinear residual differentiation
    \item Ideal for Monte Carlo
    \item Similar framework properties to matrix-free methods
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{FANM Nonlinear Iteration Results}

  \begin{table}[h!]
    \begin{center}
      \begin{tabular}{lccc}\hline\hline
        \multicolumn{1}{l}{Benchmark}& 
        \multicolumn{1}{c}{NK}&
        \multicolumn{1}{c}{FANM}&
        \multicolumn{1}{c}{NR}\\
        \hline
        Convection, Ra=\sn{1}{3} & 5 & 5 & 5\\
        Convection, Ra=\sn{1}{4} & 7 & 7 & 7\\
        Convection, Ra=\sn{1}{5} & 9 & 10 & 9\\
        Convection, Ra=\sn{1}{6} & 11 & 11 & 11\\
        \hline
        Lid Driven, Re=100 & 6 & 6 & 7\\
        Lid Driven, Re=300 & 9 & 9 & 9\\
        Lid Driven, Re=500 & 11 & 11 & 11\\
        Lid Driven, Re=700 & 14 & 10 & 12\\
        \hline
        Backward Step, Re=200 & 10 & 9 & 10\\
        Backward Step, Re=300 & 15 & 14 & 15\\
        Backward Step, Re=400 & 10 & 10 & 10\\
        Backward Step, Re=500 & 19 & 20 & 21\\
        %%
        \hline\hline
      \end{tabular}`
    \end{center}
    \caption{\textbf{Navier-Stokes benchmark comparison for nonlinear
        iterations.} \textit{Over all benchmarks, FANM performed better
        in terms of nonlinear iterations for 1 more case than the
        Newton-Krylov method.}}
    \label{tab:benchmark_nonlinear_comparison}

  \end{table}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{FANM Linear Solver Iteration Results}

  \begin{table}[h!]
    \begin{center}
      \begin{tabular}{lccc}\hline\hline
        \multicolumn{1}{l}{Benchmark}& 
        \multicolumn{1}{c}{GMRES}&
        \multicolumn{1}{c}{MCSA}&
        \multicolumn{1}{c}{Richardson}\\
        \hline
        Convection, Ra=\sn{1}{3} & 32 & 18 & 38\\
        Convection, Ra=\sn{1}{4} & 23 & 17 & 34\\
        Convection, Ra=\sn{1}{5} & 25 & 20 & 34\\
        Convection, Ra=\sn{1}{6} & 39 & 25 & 48\\
        \hline
        Lid Driven, Re=100 & 27 & 42 & 151\\
        Lid Driven, Re=300 & 35 & 52 & 133\\
        Lid Driven, Re=500 & 41 & 56 & 154\\
        Lid Driven, Re=700 & 21 & 14 & 32\\
        \hline
        Backward Step, Re=200 & 24 & 13 & 21\\
        Backward Step, Re=300 & 23 & 17 & 21\\
        Backward Step, Re=400 & 18 & 12 & 14\\
        Backward Step, Re=500 & 30 & 52 & 98\\
        %%
        \hline\hline
      \end{tabular}
    \end{center}
    \caption{\textbf{Navier-Stokes benchmark comparison for total linear
        solver iterations.} \textit{Over all benchmarks, FANM performed
        better in terms of linear solver iterations for twice as many
        cases as the Newton-Krylov method.}}
    \label{tab:benchmark_linear_comparison}
  \end{table}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{FANM CPU Time Results}

\begin{table}[h!]
  \begin{center}
    \begin{tabular}{lc}\hline\hline
      \multicolumn{1}{l}{Benchmark}& 
      \multicolumn{1}{c}{NK Speedup}\\
      \hline
      Convection, Ra=\sn{1}{3} & 338 \\
      Convection, Ra=\sn{1}{4} & 336 \\
      Convection, Ra=\sn{1}{5} & 346 \\
      Convection, Ra=\sn{1}{6} & 465 \\
      \hline
      Lid Driven, Re=100 & 299 \\
      Lid Driven, Re=300 & 322 \\
      Lid Driven, Re=500 & 288 \\
      Lid Driven, Re=700 & 488 \\
      \hline
      Backward Step, Re=200 & 400 \\
      Backward Step, Re=300 & 593 \\
      Backward Step, Re=400 & 825 \\
      Backward Step, Re=500 & 1057 \\
      %%
      \hline\hline
    \end{tabular}
  \end{center}
  \caption{\textbf{Newton-Krylov speedup over FANM.}  \textit{For all
      benchmarks the explicit MCSA preconditioning strategy caused
      significantly larger CPU times for FANM when compared to the
      Newton-Krylov solutions.}}
  \label{tab:benchmark_time_comparison}
\end{table}

>>>>

  \begin{itemize}
  \item Inject MCSA as the solver at each eigenvalue iteration
    \medskip
  \item Transport operator is static - one time cost for MCSA setup
    \medskip
  \item Current Exnihilo implementation gives the full operator
  \end{itemize}

@@

      \begin{figure}[htpb!]
        \begin{center}
          \includegraphics[width=2.25in]{solver_time.pdf}
        \end{center}
      \end{figure}

@@

      \begin{itemize}
        {\small
        \item Comparison to Trilinos Aztec Krylov solvers with ILUT
          \medskip
        \item MCLS verified for the k-eigenvalue and neutron flux in
          all groups
          \medskip
        \item MCSA converged in fewer iterations than GMRES, more
          iterations than BiCGStab
          \medskip
        \item Explicit preconditioning strategy destroys sparsity and
          elevates CPU times (Ifpack ILUT)
          \medskip
        \item Spectral radius and memory limitations combine to
          prevent solutions at finer discretizations
        }
      \end{itemize}

@@

  \begin{itemize}
  \item Sequence of Navier-Stokes benchmarks
    \begin{itemize}
    \item Thermal convection cavity problem (De Vahl Davis, 1983)
    \item Lid driven cavity problem (Ghia et al., 1982)
    \item Backward-Facing step problem (Gartling, 1990)
    \end{itemize}
    \medskip
  \item Tuning benchmark parameters varies the strength of
    nonlinearities
  \end{itemize}

@@

  \begin{itemize}
  \item Need a nonlinear solution scheme for the Navier-Stokes equations
    \bigskip
  \item Significant research on Newton methods since the 1980's
    \bigskip
  \item Newton methods often leverage Krylov solvers
    \begin{itemize}
    \item Simple implementation
    \item No operator required
    \end{itemize}
    \bigskip
  \item Monte Carlo methods need the full operator
    \bigskip
  \item Automatic construction of the linear model is available
    \begin{itemize}
    \item Operator overloading for nonlinear residual differentiation
    \item Ideal for Monte Carlo
    \item Similar framework properties to matrix-free methods
    \end{itemize}
  \end{itemize}

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Thermal Convection Cavity Results}

  \begin{columns}
    \begin{column}{0.6\textwidth}

      \small{
        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{NK}&
              \multicolumn{1}{c}{FANM}&
              \multicolumn{1}{c}{NR}\\
              \hline
              Ra=\sn{1}{3} & 5 & 5 & 5\\
              Ra=\sn{1}{4} & 7 & 7 & 7\\
              Ra=\sn{1}{5} & 9 & 10 & 9\\
              Ra=\sn{1}{6} & 11 & 11 & 11\\
              %%
              \hline\hline
            \end{tabular}`
          \end{center}
          \caption{Nonlinear iterations.}
        \end{table}

        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{GMRES}&
              \multicolumn{1}{c}{MCSA}&
              \multicolumn{1}{c}{Richardson}\\
              \hline
              Ra=\sn{1}{3} & 32 & 18 & 38\\
              Ra=\sn{1}{4} & 23 & 17 & 34\\
              Ra=\sn{1}{5} & 25 & 20 & 34\\
              Ra=\sn{1}{6} & 39 & 25 & 48\\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Total linear solver iterations.}
        \end{table}
      }

    \begin{column}{0.4\textwidth}

      \begin{itemize}
      \item Weaker preconditioning at \sn{1}{-5} case
      \item Forcing term method adapts to nonlinear convergence
      \item Constant forcing term for \sn{1}{-6} case
      \end{itemize}

      \tiny{
        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{NK Speedup}\\
              \hline
              Ra=\sn{1}{3} & 338 \\
              Ra=\sn{1}{4} & 336 \\
              Ra=\sn{1}{5} & 346 \\
              Ra=\sn{1}{6} & 465 \\
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Newton-Krylov speedup over FANM.}
        \end{table}
      }

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Lid Driven Cavity Results}

  \begin{columns}
    \begin{column}{0.6\textwidth}

      \small{
        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{NK}&
              \multicolumn{1}{c}{FANM}&
              \multicolumn{1}{c}{NR}\\
              \hline
              Re=100 & 6 & 6 & 7\\
              Re=300 & 9 & 9 & 9\\
              Re=500 & 11 & 11 & 11\\
              Re=700 & 14 & 10 & 12\\
              %%
              \hline\hline
            \end{tabular}`
          \end{center}
          \caption{Nonlinear iterations.}
        \end{table}

        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{GMRES}&
              \multicolumn{1}{c}{MCSA}&
              \multicolumn{1}{c}{Richardson}\\
              \hline
              Re=100 & 27 & 42 & 151\\
              Re=300 & 35 & 52 & 133\\
              Re=500 & 41 & 56 & 154\\
              Re=700 & 21 & 14 & 32\\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Total linear solver iterations.}
        \end{table}
      }

    \begin{column}{0.4\textwidth}

      \begin{itemize}
      \item Forcing term selection can significantly modify nonlinear
        convergence
        \medskip
      \item $\eta_k = \gamma \Big(
        \frac{||\ve{F}(\ve{u}_k)||}{||\ve{F}(\ve{u}_{k-1})||}
        \Big)^{\alpha}$
      \end{itemize}

      \tiny{
        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{NK Speedup}\\
              \hline
              Re=100 & 299 \\
              Re=300 & 322 \\
              Re=500 & 288 \\
              Re=700 & 488 \\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Newton-Krylov speedup over FANM.}
        \end{table}
      }

@@

%%---------------------------------------------------------------------------%%
\begin{frame}{Backward Facing Step Results}

  \begin{columns}
    \begin{column}{0.6\textwidth}

      \small{
        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{NK}&
              \multicolumn{1}{c}{FANM}&
              \multicolumn{1}{c}{NR}\\
              \hline
              Re=200 & 10 & 9 & 10\\
              Re=300 & 15 & 14 & 15\\
              Re=400 & 10 & 10 & 10\\
              Re=500 & 19 & 20 & 21\\
              %%
              \hline\hline
            \end{tabular}`
          \end{center}
          \caption{Nonlinear iterations.}
        \end{table}

        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{GMRES}&
              \multicolumn{1}{c}{MCSA}&
              \multicolumn{1}{c}{Richardson}\\
              \hline
              Re=200 & 24 & 13 & 21\\
              Re=300 & 23 & 17 & 21\\
              Re=400 & 18 & 12 & 14\\
              Re=500 & 30 & 52 & 98\\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Total linear solver iterations.}
        \end{table}
      }

    \begin{column}{0.4\textwidth}

      \begin{itemize}
      \item Significantly more ill-conditioned than cavity problems
      \item Forcing terms are not the culprit
      \item Multigrid is doing significantly more work
      \end{itemize}

      \tiny{
        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{NK Speedup}\\
              \hline
              Re=200 & 400 \\
              Re=300 & 593 \\
              Re=400 & 825 \\
              Re=500 & 1057 \\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Newton-Krylov speedup over FANM.}
        \end{table}
      }

%%---------------------------------------------------------------------------%%
\begin{frame}{Iterative Performance Summary}

  {\tiny

        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lcc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{Newton-Krylov}&
              \multicolumn{1}{c}{FANM}\\
              \hline
              Convection, Ra=\sn{1}{3} & $\times$ & $\times$ \\
              Convection, Ra=\sn{1}{4} & $\times$ & $\times$ \\
              Convection, Ra=\sn{1}{5} & $\times$ & \\
              Convection, Ra=\sn{1}{6} & $\times$ & $\times$ \\
              Lid Driven, Re=100 & $\times$ & $\times$ \\
              Lid Driven, Re=300 & $\times$ & $\times$ \\
              Lid Driven, Re=500 & $\times$ & $\times$ \\
              Lid Driven, Re=700 & & $\times$ \\
              Backward Step, Re=200 & & $\times$ \\
              Backward Step, Re=300 & & $\times$ \\
              Backward Step, Re=400 & $\times$ & $\times$ \\
              Backward Step, Re=500 & $\times$ & \\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Navier-Stokes benchmark comparison for nonlinear
            iterations.}
        \end{table}

        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lcc}\hline\hline
              \multicolumn{1}{l}{Benchmark}& 
              \multicolumn{1}{c}{Newton-Krylov}&
              \multicolumn{1}{c}{FANM}\\
              \hline
              Convection, Ra=\sn{1}{3} & & $\times$ \\
              Convection, Ra=\sn{1}{4} & & $\times$ \\
              Convection, Ra=\sn{1}{5} & & $\times$ \\
              Convection, Ra=\sn{1}{6} & & $\times$ \\
              Lid Driven, Re=100 & $\times$ & \\
              Lid Driven, Re=300 & $\times$ & \\
              Lid Driven, Re=500 & $\times$ & \\
              Lid Driven, Re=700 & & $\times$ \\
              Backward Step, Re=200 & & $\times$ \\
              Backward Step, Re=300 & & $\times$ \\
              Backward Step, Re=400 & & $\times$ \\
              Backward Step, Re=500 & $\times$ & \\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Navier-Stokes benchmark comparison for total
            linear solver iterations.}
        \end{table}

      \end{column}
    \end{columns}
  }

  {\small
    \begin{itemize}
    \item Over all benchmarks, FANM performed better in terms of nonlinear
      iterations for 1 more case than the Newton-Krylov method
      \medskip
    \item Over all benchmarks, FANM performed better in terms of linear
      solver iterations for twice as many cases as the Newton-Krylov
      method
    \end{itemize}
  }
